Poster figures

### 2024-11-12 analysis/poster_aarc/poster_figures.Rmd
Adds poster figures

---
title: "Untitled"
format: gfm
execute: 
  echo: false
  warning: false
  message: false
  cache: false
filters:
  - ../../_extensions/pandoc-ext/abstract-section/abstract-section.lua
  - ../../_extensions/mloubout/critic-markup/critic-markup.lua
  - ../../_extensions/ute/search-replace/search-replace.lua
search-replace:
  +S2: ACCESS-S2
  +S1: ACCESS-S1
  +CDR: NOAA/NSIDC CDRV4
bibliography: references.bib
editor: 
  markdown: 
    wrap: sentence
---

@@

cdo_options_set("-L")
source(here::here("R/functions.R"))
source(here::here("R/datasets.R"))
source(here::here("R/ggplot.R"))
```

# Abstract

balbalbal

# Introduction

Unlike Arctic sea ice, which has been steadily retreating at least since the start of satellite records in the early 80s, Antarctic sea ice had a slightly increasing trend that models systematically failed to reproduce and puzzled researchers [@hobbs2016]. 
Then, in 2016 Antarctic sea ice extent dropped precipitously and has been at low and record low levels since, highlighting even more our lack of understanding of Antarctic sea ice variability and change. [ref: that CESM run with forced winds].

Antarctic sea ice is in a remote location but the potential impacts of its variability and trends extend far into the lower latitudes, affecting ocean and atmospheric circulation, ocean carbon uptake and biological processes. 
It's likely that a reduction in sea ice extent would reduce the temperature gradient between high and low latitudes (and vice versa), which in turn influences the strength and location of the eddy-driven jet, which is a crucial component of weather events. 
For the northern hemisphere, it has been hypothesised that the reduction in sea ice in the northern hemisphere has lead to a weaker, wavier jet, increasing the frequency of extreme events [@barnes2015]. 
There is some evidence of this effect in the Southern Hemisphere. 
Models show a weakened vortex in the middle atmosphere, increased vertical wave flux and greater zonal wave 1 amplitude in Springs following a low sea ice maximum. 
But models don't fully agree on the exact phase of the zonal wave response and observations don't show a clear signal [@rea2024]. 
The jet response to sea ice loss is also very model dependent, particularly sensitive to its the climatological location [@ayres2019]. 
It is also small compared with the direct effect of CO2 increase and Sea Surface Temperature warming [@ayres2019].
On climate scales, it is thought that the signal of a reduction in Antarctic sea ice is like a "mini global warming" and experiments also predict a weakening and northerly shift of the eddy-driven [@raphael2011; @england2018; @england2020; @ayres2022].

Correctly modelling Antarctic sea ice is not only necessary for process understanding and climate projections to inform adaptation strategies. 
Accurate seasonal to subseasonal forecast are crucial for operations in and around the Antarctic continent [@desilva2020 @wagner2020]. 
However, Antarctic sea ice forecasts have been challenging both due to model biases and inherent large variability and complexity and it has lagged behind Arctic sea ice forecasts [@zampieri2019]. 
Dynamical seasonal forecasts of Summer Antarctic sea ice have been shown to perform worse than relatively simple statistical methods [@massonnet2023], which also underscores the need for better understanding of sea ice dynamics.

{==**S2** Introduce S2==}

{==**Objective.** Evaluate S2.==}

# Data and methods

## +S2

+S2 [@wedd2022] is the Australian Bureau of Meteorology's current seasonal prediction system that replaced +S1 [@hudson2017] in October 2021. 
Both +S2 and +S1 consist on the Global Atmosphere 6.0 (GA6) [@williams2015; @waters2017], Global Land 6.0  [@best2011; @waters2017], Global Ocean 5.0 [@madec2013; @megann2014] and Global Sea Ice 6.0 [CICE; @rae2015].
The atmosphere has a N216 horizontal resolution(~60km in the mid-latitudes) with 85 levels. 
The land model uses the same horizontal grid with 4 soil levels. 
The ocean component has a 1/4º resolution with 75 vertical levels. 
The ice component --based on CICE version 4.1-- has the same resolution than the ocean and 5 sea ice categories as well as an open water category.

Both systems took atmospheric initial conditions derived from ERA-interim [@dee2011] for the hindcast and from the Bureau's operational analysis for real-time forecasts. 
The main difference between the two are the ocean and ice initial conditions. 
+S1 initial conditions came from from the Met Office FOAM system, which uses a multivariate, incremental three-dimensional variational (3D-Var), first-guess-at-appropriate-time (FGAT) DA scheme [@waters2015] and assimilates sea surface temperature (SST), sea surface height (SSH), in situ temperature and salinity profiles, and sea ice concentration. 
+S2, instead, runs from initial conditions generated by the in-house data assimilation scheme described in @wedd2022. 
This scheme is a weakly coupled ensemble optimal interpolation method and assimilates temperature and salinity profiles from EN4 [@good2013] for the hindcast and from the WMO Global Telecommunication System (GTS) and both the Coriolis and USGODAE Global Data Assembly Centers (GDACs) for the real-time forecast.
SSTs are nudged to Reynolds OISSTv2.1 [@reynolds2007] for the hindcast and to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and Beggs 2008) for the real-time forecast in areas where SSTs are larger than 0ºC.
Relevant for this work, sea ice concentrations are not assimilated.

For evaluation we use hindcast for the period 1981--2023{\>\>Check\<\<}.
Anomalies will be taken with respect to the 1981--2011 climatology computed from the reanalysis.
Climatology is smoothed with a 11 day running mean.

## Verification datasets

@@

There is not a lot of data on sea ice properties, especially for things like thickness, age, etc.
However there are relatively reliable satellite-derived estimates of sea ice concentration, which estimates the proportion of each grid area that is covered with ice.

These data are not perfect. \cite{meier2019}
Sensor sensitivity and resolution. 
Weather filtering.
Land 
To account for observational uncertainty we performed the verification on other data sources. 
We find that the spread among the different datasets is minimal compared with the difference with the +S2 and +S1 forecasts, so the conclusions are independent of the dataset used.

We use NOAA/NSIDC's Climate Data Record V4 [CDR; @meier2014] as the primary verification dataset. 
Its sea ice concentration data are estimated taking the maximum value of the NASA Team [ref] and NASA Bootstrap [@comiso2023] algorithms to reduce their low concentration bias [@meier2014; @meier2021].
Both algorithms use data from the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7 satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and -F13 satellites. 
The data has a resolution of 25 by 25km and daily from 1978 onwards.

Results based on two other different datasets are provided as supplementary material:

1. The NASA Bootstrap [@comiso2023] product is... it is not an independent dataset from NOAA/NSIDC CDRV4.

1. The European Organisation for the Exploitation of Meteorological Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility (OSI SAF) [ref] is another satellite-derived sea ice concentration product based on mostly the same sensors as the NOA CDR but computed independently using different algorithms.

## Error measures

For evaluation purposes, we use a series of measures.

### Sea ice extent

Sea Ice Extent is defined as the area of ocean covered with at least 15% ice.
This threshold is motivated by the limitations in satellite retrieval, which is increasingly unreliable for lower sea ice conditions.

Sea Ice Extent is a rough global measure, but a model could have relatively accurate extent of ice but with different distributions.
We use two other measures to account for these errors.

We compute Root Mean Square Error of sea ice concentration anomalies.

We also compute the Integrated Ice Edge Error (IIEE) [@goessling2016].
This is defined as the area in which the model misspredicts sea ice concentration being above or belong 15% ice.
That is, dichotomise sea ice concentration into areas with more and less than 15% sea ice both in the forecast and observations; the IIEE is the area in which forecast and observations differ.

# Results and discussion

## Reanalysis

### Bias

@@

```{r fig-hindcast-extent}
#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the median sea ice extent of +CDR. 
hindcast_extent |> 
  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                              lag = as.numeric(as.Date(time) - forecast_time),
                              model)] |>
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
  # _[]
  ggplot(aes(time2, mid)) +
  geom_line(data = extent_climatology[dataset == which_dataset], 
            aes(x = as.Date(time),
                group = dataset)) + 
  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +

  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)], 
             aes(fill = month), 
             shape = 24, size = 2.2) +

  geom_point(data = \(x) x[lag == 1], aes(fill = month), 
             shape = 21, size = 2) +

@@

@fig-hindcast-extent a shows the seasonal cycle of Sea Ice Extent for the +S2 hindcast and +CDR
+S2 shows a severe low extent bias, especially in the late summer-early autumn.
This is due primarily to a faster and longer melt season between January and March and slower growth during March and April This is then balanced with faster growth between May and July ([Fig. @fig-mean-growth]).
Many S2S systems exhibit this systematic underestimation during the sea ice minimum and early freezing season [@massonnet2023] and could indicate problems in the representation of thermodynamics in the model [@zampieri2019].

Comparing +S2 with +S1, the latter has a smaller bias, especially at low lags ([Fig. @fig-hindcast-extent b]) even though the typical growth rates are very similar between both models ([Fig. @fig-mean-growth]).
At larger lags, +S1's bias in summer and autumn is very similar to +S2's.
This suggests that this lower sea ice state is closer to the models' equilibrium, indicating that it is an issue with model formulation that was being corrected by the data assimilation system in +S1.

At long lags, sea ice extent loses most of the initial condition memory and reverts to the model's preferred equilibrium state.
Therefore we can estimate the latter using the hindcasts with the largest possible lag, which is shown in triangles in @fig-hindcast-extent for the same dates as the initial conditions.
The difference between the two is the effect of the data assimilation.

The equilibrium of +S1 and +S2 is very similar (comparing the triangles in each panel in [Fig. @fig-hindcast-extent]), owing to both having the same model formulation.
From June to October, in +S2 circles move away from triangles and towards observations, indicating that the information from the ocean and atmosphere data assimilation is getting to sea ice and affecting the initial conditions.
The rest of the year, there is little if any difference between circles and triangles in +S2, indicating that almost no data assimilation is taking place and the sea ice component of the model is virtually free-running.

```{r fig-mean-growth}
#| fig-cap: Median daily sea ice extent growth of +S1 and +S2 hindcasts and observations. Values are smoothed with a 2-degree loess smooth with a 30 day window.

dif_hindcast <- hindcast_extent |> 
  copy() |> 
  _[order(time)] |> 
  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(model, forecast_time, member)] |> 
  _[, median_ci(dif), by = .(model, time = update(time, year = 2000))]

extent_daily |> 
  _[order(time)] |> 
  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
  _[dataset == which_dataset] |> 
  ggplot(aes(time, mid)) +
  geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
  geom_smooth(aes(group = dataset, color = dataset),
              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
  geom_smooth(data = dif_hindcast,
              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
  scale_x_datetime(NULL, date_breaks = "1 month", 
                   date_labels = "%b", expand = c(0, 0)) +
  scale_y_continuous(NULL, labels = labels_extent) +
  scale_color_models
```

@@

```{r fig-bias}
#| fig-width: 9
#| fig-height: 6
#| fig-cap: +S2 reanalysis sea ice concentration bias compared with NSIDC sea ice concentration.

climatologies |> 
  _[model %in% c("S2", which_dataset)] |> 
  dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
  setnames(which_dataset, "obs") |> 
  ggplot(aes(xgrid, ygrid)) +
  geom_contour_fill(aes(z = S2 - obs, fill = after_stat(level)), 
                    breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
  scale_fill_divergent_discretised(paste0("Bias with respect to ",
                                          labels_models[which_dataset]),
                                   low = scales::muted("red"),
                                   high = scales::muted("blue")) +
  geomcoord_antarctica +
  geom_antarctica_fill +
  facet_wrap(~ lubridate::month(time, label = TRUE)) +
  wide_legend +
  tag_facets()
```

@@

### Anomalies

{==intro connecting subsections==}

```{r fig-extent-anom}
#| fig-cap: Sea ice extent anomalies for +S1 and +S2 (black) and +CDR (blue).

months_difference <- function(x, y) {
  lubridate::interval(y, x) %/% months(1)
}

monthly_extent <- extent_daily |> 
  _[aice == 0, aice := NA]  |> 
  _[, .(aice = mean(aice)), by = .(dataset, time = as.Date(round_date(time, "month")))] |>
  _[, aice := Anomaly(aice, year(time) %between% c(1981, 2011), na.rm = TRUE), 
    by = .(month(time), dataset)]

N <- uniqueN(monthly_extent$time)

hindcast_extent |> 
  copy() |> 
  # _[model == "S2"] |> 
  _[, aice := aice - mean(aice[year(time) %between% c(1981, 2011)]), 
    by = .(lag, model, time = update(time, year = 2000))] |> 
  _[, .(aice = mean(aice)), by = .(model, forecast_time, time = as.Date(round_date(time, "month")))] |> 
  _[, lag := months_difference(time, forecast_time) + 1] |> 
  _[lag %in% c(1, 3, 5, 7)] |> 
  ggplot(aes(time, aice)) +
  geom_line(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
              _[dataset == which_dataset],
            aes(group = dataset, color = dataset)) +
  # geom_smooth(data = monthly_extent[year(time) <= max(year(hindcast_extent$time))] |> 
  #               _[dataset == "cdr"], 
  #             aes(group = dataset, color = dataset),
  #             se = FALSE, span = 10*12/N, n = 366) +
  geom_line(aes(color = model))  +
  # geom_smooth(aes(color = model), se = FALSE, span = 10*12/N, n = 366) +
  scale_y_continuous(NULL,
                     labels = labels_extent) +
  scale_x_date(NULL, expand = c(0, 0)) +
  facet_grid(lag ~ model, labeller = labeller(lag = \(x) paste0("Lag: ", x))) +
  scale_color_models
```

@fig-extent-anom shows monthly sea ice extent anomalies forecasted at selected lags.
The anomalies in this case were computed with respect of the climatology of each lag, which is a way of bias-correction.
Compared with +S1, +S2 anomaly forecast is relatively poor even in the first month, which stays relatively skilful even at lag 3.
+S2 shows much bigger variability than observations, with dramatic lows between 1995 and 2007 and highs between 2007 and 2015.

@@

```{r fig-rmse}
#| fig-cap: Median and 95% coverage of sea ice concentration anomalies RMSE as a function of forecast lag for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies. 
#| fig.height: 5
#| fig.width: 9

@@

  geom_ribbon(aes(ymin = low, 
                  ymax = high, 
                  color = version, 
                  fill = version,
                  group = interaction(obs_dataset, version)),
              alpha = 0.1) +
  geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +
  scale_x_continuous(breaks = NULL) +
  labs(y = "RMSE", 
       x = "Lag") +
  scale_color_models +
  scale_fill_models +
  #   scale_color_models +
  #   scale_fill_models +
  facet_wrap(~ month, labeller = labeller(month = labels_month)) 
```

@fig-rmse shows the median and 95% range of RMSE of sea ice concentration anomalies for +S2 forecasts compared with a benchmark of persistence.
This figure uses +CDR data, but the results are nearly identical compared with ERA5 or Bootstrap.
Due to errors in the initial conditions, it is expected that a persistence forecast would be better than the model forecast at very short lags, but that the persistence forecast errors would grow faster and eventually surpass the mode forecast, at which time is statistically useful {==I've got this from CC at the ICTP summer school and makes sense, but it would be great to have a reference?=
=}.
Here the persistence errors are almost always lower than the +S2 forecast, indicating that the model doesn't have skill at any lag and in any month.
The only exception is the RMSE around February forecasted from June onwards.

@@

```{r fig-iiee}
#| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lag for all forecast initialised on the first of each month for +S1 and +S2 hindcasts. For each month, the number in parenthesis indicates the minimum lag at which the the mean error of each model is not statistically significant at a 1% level using a two-sisded t-test.

@@

  scale_y_continuous(labels = labels_extent) +
  scale_x_continuous(breaks = NULL) +
  scale_color_models +
  scale_fill_models +
  labs(y = NULL, 
       x = "Lag") +
  #   scale_color_models +
  #   scale_fill_models +
  facet_wrap(~ month, labeller = labeller(month = labels_month_significant)) 
```

To compare +S2 with +S1, we computed the IIEE for both models.
This error measure is shown in @fig-iiee for all lags and forecasts initialised at the first of every month.
+S1 has lower error at short lags at all months, with the errors converging as the forecast goes on.
The time to convergence depends on the month and it can be as fas as two weeks June and July to as large as 160 days for forecasts initialised in February. 
Since the only difference between these forecasts are the initial conditions, this timescale is an indication of the the memory of sea ice to initial conditions; at least from October to March when the data assimilated form the other components has little to no influence on sea ice.

Also evident in @fig-iiee is the difference in the error spread at short lags between +S2 and +S1.
In all month +S1 shows a small error spread at lag 1, indicating that the error in the initial conditions not only is small, but also fairly constant.
This spread then grows towards a climatological spread as errors accumulate differently in different forecasts.
For +S2, this is true only only between July and October, approximately.
For all other months, the error spread is more or less stable throughout the forecast window, indicating that not only the initial error is high, but it's not constant.

```{r fig-iiee-variance}
#| fig-cap: Mean spread of IIEE at different lags for different models. 
#| fig.height: 9
#| fig.width: 4
labels_extent_2 <- function(x) {
  m <- which.max(x)
  x <- scales::label_number(scale = (1e-12)^2)(x)
  x[m] <- paste0(x[m], "\nM km⁴")
  x
}

rbind(mean_variance = errors |>
        _[measure == "iiee"] |>
        _[obs_dataset == which_dataset] |>
        _[, var(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
        _[, .(value = mean(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],

      variance_of_mean = errors |>
        _[measure == "iiee"] |>
        _[obs_dataset == which_dataset] |>
        _[, mean(value, na.rm = TRUE), by = .(version, lag, time_forecast)] |> 
        _[, .(value = var(V1, na.rm = TRUE)), by = .(version, month(time_forecast), lag)],
      idcol = "component") |>

  ggplot(aes(lag, value)) +

  geom_line(aes(color = version), linewidth = 1) +
  scale_y_continuous("Variance", 
                     labels = labels_extent_2) +
  # scale_color_models +
  scale_x_continuous("Lag", breaks = NULL) +
  scale_color_models +
  facet_grid(month ~ component, 
             labeller = labeller(month = labels_month, 
                                 component = c(mean_variance = "Mean variance",
                                               variance_of_mean = "Variance of mean"))) 
```

The large initial error spread could be due either to large spread of ensemble members or due to a large spread of individual forecasts.
@fig-iiee-variance splits the IIEE variance for each lag into the mean variance of each individual forecast and the variance of the mean error of each individual forecast, which adds up to the total variance.
The average variance of each forecast is almost identical between forecast systems in all months.
This shows that the ensemble spread of individual forecasts evolves identically, which, again, it's not unexpected because both systems share the same model formulation.
This also shows that the perturbation scheme in +S2 is comparable to the one in +S1.

On the other hand, the spread of the mean error is always larger in +S2 than +S1.
The difference is particularly large at short lags in some months, which coindice with the ones in which the data assimilation scheme is not influencing sea ice initial conditions.

## Conclusions

# References

::: {#refs}
:::

# Suplementary figures

<!-- ## ERA5 -->

<!-- ```{r dataset_era} -->
<!-- which_dataset <- "era5" -->
<!-- ``` -->

<!-- ```{r fig-hindcast-extent-era5} -->
<!-- which_dataset <- "era5" -->
<!-- <<fig-hindcast-extent>> -->
<!-- ``` -->

<!-- ```{r fig-hindcast-extent-bt} -->
<!-- which_dataset <- "bt" -->
<!-- <<fig-hindcast-extent>> -->
<!-- ``` -->

<!-- ```{r fig-mean-growth} -->
<!-- <<fig-mean-growth>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<fig-bias>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<<fig-extent-anom>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<fig-rmse>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<fig-iiee>> -->
<!-- ``` -->

<!-- ```{r} -->
<!-- <<fig-iiee-variance>> -->
<!-- ``` -->

>>>>

---
title: "Poster figures"
output: 
  html_document:
    self_contained: false  
date: "2024-11-11"
---

@@

cdo_options_set("-L")
source(here::here("R/functions.R"))
source(here::here("R/datasets.R"))
source(here::here("R/ggplot.R"))

```{r}

my_theme <- function(base_size = 20) {
  ggplot2::theme_minimal(base_size = base_size) +
            ggplot2::theme(panel.background = ggplot2::element_rect(fill = "#fafafa", color = NA),
                  legend.position = "bottom",
                  legend.title.position = "top", 
                  legend.title = ggplot2::element_text(hjust = 0.5),
                  legend.frame = ggplot2::element_rect(color = "black", linewidth = 0.4),
                  legend.key.height = grid::unit(0.75, "lines"),
                  axis.text.y = element_text_first_line())
}

ggplot2::theme_set(my_theme())

@@

@@

```{r fig-hindcast-extent, fig.width=2306/150, fig.height=1352/150}
#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the median sea ice extent of +CDR. 
hindcast_extent |> 
  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                              lag = as.numeric(as.Date(time) - forecast_time),
                              model)] |>
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
  # _[]
  ggplot(aes(time2, mid)) +
  geom_line(data = extent_climatology[dataset == which_dataset], 
            aes(x = as.Date(time),
                group = dataset),
            linewidth = 1) + 
  geom_line(aes(color = month, group = interaction(forecast_time, time != time2)),
            linewidth = 1) +

  geom_point(data = \(x) x[mday(time) == 2, .SD[which.max(lag)], by = .(forecast_time, model)], 
             aes(fill = month), 
             shape = 24, size = 4) +

  geom_point(data = \(x) x[lag == 1], aes(fill = month), 
             shape = 21, size = 4) +

@@

@@

```{r hindcast_clim}
zeropad <- function(x, n = 2) formatC(x, width = 2, flag = "0")
monnb <- function(d) { lt <- as.POSIXlt(as.Date(d, origin="1900-01-01")); 
lt$year*12 + lt$mon } 
mondf <- function(d1, d2) { monnb(d2) - monnb(d1) }

months <- 1:12
hindcast_clim <- lapply(setNames(months, months), \(month) {
  here::here(glue::glue("data/derived/hindcast_climatology/mo_aice_1981{zeropad(month)}01_mean.nc")) |> 
    ReadNetCDF("aice")
}) |> 
  rbindlist(idcol = "forecast_month") |> 
  _[, forecast_month := as.numeric(forecast_month)] |> 
  _[, forecast_time := lubridate::make_datetime(1981, forecast_month, 01)] |> 
  _[, time := lubridate::floor_date(time, "month")] |> 
  _[, lag := mondf(forecast_time, time)] |> 
  _[lag == 0] |> 
  _[, time2 := update(time, year = 2000)]

obs_clim <- CDR() |> 
  climatology() |> 
  ReadNetCDF(c(obs = "aice")) |> 
  setnames("time", "time2")

hindcast_clim <- hindcast_clim |> 
  merge(obs_clim) |> 
  _[, month := month(time)] 
```

```{r fig-bias, fig.height=2306/150, fig.width=2306/150}
#| fig-cap: 1-month lag +S2 forecast bias compared with NSIDC.

hindcast_clim |> 
  # _[time == time[1]] |> 
  ggplot(aes(xgrid, ygrid)) +
  geom_contour_fill(aes(z = aice - obs, fill = after_stat(level)), 
                    breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
  scale_fill_divergent_discretised(NULL,
                                   low = scales::muted("red"),
                                   high = scales::muted("blue")) +
  geomcoord_antarctica +
  geom_antarctica_fill +
  facet_wrap(~ lubridate::month(time, label = TRUE)) +
  wide_legend +
  tag_facets() +
  theme(panel.ontop = TRUE, 
        panel.background = element_blank(),
        panel.grid = element_line(colour = alpha("gray50", 0.5), linewidth = 0.4))
```

@@

@@

```{r fig-rmse, fig.height=1524/150, fig.width=2306/150}
#| fig-cap: Median and 95% coverage of sea ice concentration anomalies RMSE as a function of forecast lag for all forecast initialised on the first of each month compared with a reference forecast of persistence of anomalies. 
#| fig.height: 5
#| fig.width: 9

@@

  geom_ribbon(aes(ymin = low, 
                  ymax = high, 
                  color = version, 
                  fill = version,
                  group = interaction(obs_dataset, version)),
              alpha = 0.1) +
  geom_line(aes(color = version, group = interaction(obs_dataset, version)), linewidth = 1) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  labs(y = "RMSE", 
       x = "Lag") +
  scale_color_models +
  scale_fill_models +
  #   scale_color_models +
  #   scale_fill_models +
  facet_wrap(~ month, labeller = labeller(month = labels_month)) 
```

@@

```{r fig-iiee, fig.height=1524/150, fig.width=2306/150}
#| fig-cap: Median and 95% coverage of Integrated Ice Edge Error as a function of forecast lag for all forecast initialised on the first of each month for +S1 and +S2 hindcasts. For each month, the number in parenthesis indicates the minimum lag at which the the mean error of each model is not statistically significant at a 1% level using a two-sisded t-test.

@@

  scale_y_continuous(labels = labels_extent) +
  scale_x_continuous(breaks = NULL, expand = c(0, 0)) +
  scale_color_models +
  scale_fill_models +
  labs(y = NULL, 
       x = "Lag") +
  #   scale_color_models +
  #   scale_fill_models +
  facet_wrap(~ month, labeller = labeller(month = labels_month_significant)) 
```
